\vspace{-0.5cm} Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data\\[0.5cm] \Large Replies to Reviewer's Comments

### 2018-09-13 review/review.Rnw
Make minor corrections on the point-by-point responses

\textbf{\underline{Specific reviewer's comments}:}
\begin{enumerate}
\item \textit{There has been some active research to find a better
    approximation for normalizing constant $Z(\lambda, \nu)$ from which
    the expressions for both mean and variance will be derived. Please
    refer to the Gaunt et al. (2016) for the latest results.}\\[0.1cm]
  \textbf{\underline{Response:}} Thanks for this reference. We have
  cited the paper published in the Annals of the Institute of
  Statistical Mathematics \cite{Gaunt2017}.
\item \textit{It seems that the paper missing some latest literature on
    CMP distribution. For example, the distributional properties of CMP
    were found in Daly and Gaunt (2015).}\\[0.1cm]
  \textbf{\underline{Response:}} Thanks for this reference. We have
  cited the paper published in the ALEA, Latin American Journal of
  Probability and Mathematical Statistics \cite{Daly2016}.
\item \textit{The likelihood equation (4.1) needs more details. Can we
    derive score equations? It seems like the likelihood has become even
    more complicated. Some of the nice properties of the CMP likelihood
    such as the canonical link are lost.}\\[0.1cm]
  \textbf{\underline{Response:}} The log-likelihood function cannot be
  derived in relation ti the dispersion parameter using any
  parametrization. We obtained the Hessian matrix by central finite
  differences using the Richardson method implemented in
  \texttt{R}. Although the expression of reparametrized log-likelihood
  (Equation 3.1) seems more complicated, this function is better behaved
  than the function expressed in the original form (see Figure
  8). Consider the exponential family of distributions, with probability
  function given by
  $$f(y) = \exp\{\sigma[\theta y - b(\theta)] - c(y, \phi)\},$$
  the COM-Poisson belongs to this family with $\sigma=1$,
  $\theta = \log(\lambda)$, $b(\theta) = \log(Z(\lambda, \nu))$ and
  $c(y, \phi)=-\nu\log(y!)$, only if $\nu$ is known. Since we don't have
  $b'(\theta) = \partial b(\theta)/\partial \theta$ in closed form, the
  canonical link function has no important rule here. In this case, the
  link function maps $\theta$ in the parameter space of $\lambda$.  In
  the reparametrized version of the COM-Poisson, the log link function
  maps the parameter space and provide simple interpretations (in terms
  of rate ratios).
\item \textit{For model fitting, the authors used general purpose
    optimization function in R. Did you observe any non convergence
    situations during your numerical studies.}\\[0.1cm]
  \textbf{\underline{Response:}} Yes, we have 37 situations where the
  algorithm did not converge in the simulation study (see table
  below). This occurred only for strong overdispersion and small sample
  sizes. We reported these situations in the updated version of the paper.
  \begin{table}[H]
    \centering
    \begin{tabular}{lcccc}
      \hline
      & $\phi=-1.6$ & $\phi=-1$ & $\phi=0$ & $\phi=1.8$ \\
      \hline
      $n=50$ &    29 &  0 &   0 &   0 \\
      $n=100$ &   8  &  0 &   0 &   0 \\
      $n=300$ &   0  &  0 &   0 &   0 \\
      $n=1000$ &  0  &  0 &   0 &   0 \\
      \hline
    \end{tabular}
  \end{table}
\item \textit{Numerical derivatives with CMP tend to be very slow and
    prone to errors. It would be good to have a computation times from
    the method for different sample sizes and dispersion
    levels.}\\[0.1cm]
  \textbf{\underline{Response:}} We have computed the times to fit the
  models for the case studies in the paper, they have different
  dispersion levels and sample sizes. However, the Figure below show the
  times to fit in the simulation study. The results show that it took
  more time to fit the models under equidispersion. This result is due
  to choose of start values and normalizing constant upper bound. We
  set, for all models, the upper bound 500 and we start the algorithm
  from the real parameters.

  \hspace*{-1cm}
\begin{minipage}{\textwidth}
<<simul-times, fig.width=9, fig.height=3.5>>=
@
\end{minipage}

\item \textit{In Section 5, the simulated counts are ranging from 3 to
    27. This is too limited. Please increase the range.}\\[0.1cm]
  \textbf{\underline{Response}:} The simulation study was performed with
  expectations varying from 3 to 27. Therefore, considering the
  different scenarios ($\phi=-1.6, \phi=-1, \phi=0$ and $\phi=1.8$), we
  cover a wide range of counts. The Figure below shows the variation of
  the counts for $\mu=27$ for the different scenarios.

@@

\item \textit{From the presentation point of view, focusing more on
    applications would be good as there is not much novelty in the
    methodology.}\\[0.1cm]
  \textbf{\underline{Response:}} In this paper, we propose a
  mean-parametrization of the COM-Poisson model. This parametrization
  introduced $\mu = \lambda^{1/\nu} -(\nu-1)/(2\nu)$, a simple function
  of the original parameters $\lambda$ and $\nu$. We showed that the new
  parameter space has nice properties like i) the orthogonality between
  $\mu$ and $\phi$ (and consequently between $\beta_j$ and $\phi$ in a
  regression setting), leading to better computation and asymptotic
  (normal-based) inference \citep{Ross1970}; and ii) interpretability of
  the $\beta_j$ in terms of the rate ratios since
  $\mu_i\approx \text{E}(Y_i)$. Moreover, we characterized the
  COM-Poisson distribution in terms of the dispersion, zero-inflation
  and heavy-tail indexes that illustrate its applicability in the real
  count data. These are the methodological novelties of the paper
  besides the exploration of the results. However, in the updated
  version of the paper, we provided a better discussion of the case
  studies.
\end{enumerate}

@@

\textbf{\underline{Third paragraph}:} \textit{The authors also hint that
  $var(Y_i)\approx \mu_i \exp(-\phi)$ (see the quadratic error analysis
  leading to Fig. 1). This is curious given that
  $avar(Y_i)=\mu\exp(-\phi) + (\exp(\phi)-1)/(2\exp(2\phi))$ and the
  latter summand converges to $-\infty$ as $\phi\to-\infty$. This latter
  summand is 0 or close to 0 for $\phi=0$ or $\phi>0$. Taken together,
  these last two sentences imply that the $\mu_i\exp(-\phi)$ may be a
  reasonable approximation to the variance when there is equi- or under-
  dispersion, but it may be unreasonable when there is lots of
  overdispersion. It seems that some of the results, simulations, and
  sample analyses reflect this (e.g. Fig 3, Fig 6-8, ``deviance function
  shape under strong overdispersion $\phi=-1.6$ is not as well
  behaved...'', p. 19, etc.). This needs to be addressed.}\\[0.3cm]
\textbf{\underline{Response:}} We thanks the referee for exploring this
point. However, we didn't use the approximation of variance
($a\text{Var}(Y_i)$) to reparametrize or fit the COM-Poisson model. To
compute the variances, that is showed in the Figure 3(a) and (b) we used
$\text{Var}(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y
p(y)]^2$.
As highlighted in the Figure 1(b), the approximation is not accurate,
mainly for overdispersion ($\phi<0$). The Figure (a) and (b) below shows
the behavior of $a\text{Var}(Y_i)$ function by fixing $\mu$ and $\phi$,
respectively, illustrating the referees' statement. The $a\text{Var}(Y)$
leads to negative values for variance for small $\mu$ and $\phi<0$ that
shows that the approximation is not reasonable for this region of the
parameter space.

@@

\textbf{\underline{Fifth paragraph}:} \textit{This last point also hints
  that a quasi-Poisson model based on the assumption that
  $var(Y_i)\propto\mu_i$ (see p. 14) may not be reasonable for analyzing
  overdispersed data if the $COMPo$ model truly holds. Is this the
  case?}\\[0.1cm]
\textbf{\underline{Response:}} The quasi-Poisson model is specified by
second-moment assumptions (expectation and variance). As highlighted in
Figure 3(a) (and in referee's previous comment), the mean-variance
relationship for COM-Poisson is linear. Therefore, is suitable to
analyze data generated according to the COM-Poisson distribution, using
the assumption $\text{Var}(Y_i)\propto \mu_i$. The advantage of the
COM-Poisson approach is that it corresponds to a fully specified
probability model allowing to compute the likelihood and some useful
measures like deviance, AIC, BIC, LRT, etc.
\vspace{0.4cm}

\textbf{\underline{Sixth paragraph}:} \textit{Consider the example of
  Section 6.1. If I understand matters correctly, it is surprising that
  the $\beta$ estimates under the
  $[COMPo(\lambda_i, \nu), g(m(\lambda_i, \nu)) = \beta^\top x_i]$ model
  are so different from those under the
  $[COMPo(\mu_i, \phi), g(\mu_i) = \beta^\top x_i]$. See the results in
  Table 2, for example. I would think that if the approximation
  $m(\lambda_i, \nu) \approx \mu_i$ is reasonable, which your quadratic
  error results seem to indicate (see Fig 1), then the ML estimates of
  $\beta$ would be very similar. Exactly what models are being fitted
  here? This needs clarification.}\\[0.1cm]
\textbf{\underline{Response:}} Note that the $\lambda_i$ does not
represent the expectation, nor approximately. The quadratic errors in
Figure 1(a) is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.
The models fitted in the case studies are
$[\text{COMPo}(\lambda_i, \nu), \log(\lambda_i) = \beta^\top x_i]$ and
$[\text{COMPo}(\mu_i, \phi), \log(\mu_i) = \beta^\top x_i]$. Therefore,
the $\beta$'s estimates are comparable only if $\phi=0$ (Poisson special
case) otherwise they are on different scales. We make it clear in which
parameter the linear predictor is placed in the current version of the
paper. To understand the relationship of the coefficients, consider
$\log(\lambda_i) = \beta^o$ and $\log(\mu_i) = \beta^*$, so $\beta^*$
expressed in terms of $\beta^o$ is
$\beta^* = \exp(\beta^o-\phi) - (\exp(\phi) - 1)/(2\exp(\phi))$.
\vspace{0.4cm}

@@

\begin{enumerate}
\item \textit{Consider the constraint ``$\mu > 0$'', after equation
    (3.2). What constraints does this impose on the $\lambda$ and
    $\nu$. Is this reasonable?}\\[0.1cm]
  \textbf{\underline{Response:}} The constraint $\mu>0$ implies that
  $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figure below show (a) the
  constraint border in the parameter space; and (b) the expected values
  (obtained by $\sum y p(y)$). The constraint lies on the very small
  $\lambda$ and $\nu>1$ (underdispersion). This parameter region is
  related to small expected values (smaller than 0.1) and
  underdispersion. Therefore, although this constraint is undesirable,
  it does not prejudice the application of the model. We thank the
  referee for highlighting this point. We have discussed it in the
  updated version.

@@

\item \textit{The Shmueli approximations to the mean and variance hold
    under certain conditions. Are those conditions met in practice? For
    your examples?}\\[0.1cm]
  \textbf{\underline{Response:}} We used only the mean approximation
  given by Shmueli to reparametrize the COM-Poisson model. Moreover, as
  discussed in Figure 1(a), the the errors are close to 0 for the
  parameter grid evaluated and present no clear relation with regions
  gives by \cite{Shmueli2005} ($\phi \leq 0$ and
  $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$. Except for
  counts with small averages (close to 0), the reparametrized model fits
  well. For examples presented in the paper, the parameter estimates
  are summarizes below.

<<contraints-case-studies, results="asis", cache=TRUE>>=
@

>>>>

\textbf{\underline{Specific reviewer's comments}:}
\begin{enumerate}
\item \textit{There has been some active research to find a better
    approximation for normalizing constant $Z(\lambda, \nu)$ from which
    the expressions for both mean and variance will be derived. Please
    refer to the Gaunt et al. (2016) for the latest results.}\\[0.1cm]
  \textbf{\underline{Response:}} Thanks for this reference. We have
  cited the paper published in the Annals of the Institute of
  Statistical Mathematics \cite{Gaunt2017}.
\item \textit{It seems that the paper missing some latest literature on
    CMP distribution. For example, the distributional properties of CMP
    were found in Daly and Gaunt (2015).}\\[0.1cm]
  \textbf{\underline{Response:}} Thanks for this reference. We have
  cited the paper published in the ALEA, Latin American Journal of
  Probability and Mathematical Statistics \cite{Daly2016}.
\item \textit{The likelihood equation (4.1) needs more details. Can we
    derive score equations? It seems like the likelihood has become even
    more complicated. Some of the nice properties of the CMP likelihood
    such as the canonical link are lost.}\\[0.1cm]
  \textbf{\underline{Response:}} The log-likelihood function cannot be
  derived in relation to the dispersion parameter using any
  parametrization. We obtained the Hessian matrix by central finite
  differences using the Richardson method as implemented in R. Although
  the expression of reparametrized log-likelihood (Equation 3.1) seems
  more complicated, this function is better behaved than the function
  expressed in the original form (see Figure 8). Consider the
  exponential family of distributions, with probability function given
  by
  $$f(y) = \exp\{\sigma[\theta y - b(\theta)] - c(y, \phi)\},$$
  the COM-Poisson belongs to this family with $\sigma=1$,
  $\theta = \log(\lambda)$, $b(\theta) = \log(Z(\lambda, \nu))$ and
  $c(y, \phi)=-\nu\log(y!)$, only if $\nu$ is known. Since we don't have
  $b'(\theta) = \partial b(\theta)/\partial \theta$ in closed form, the
  canonical link function has no important role here. In this case, the
  link function maps $\theta$ in the parameter space of $\lambda$.  In
  the reparametrized version of the COM-Poisson the log link function
  maps the parameter space and provides simple interpretations (in terms
  of rate ratios).

\item \textit{For model fitting, the authors used general purpose
    optimization function in R. Did you observe any non convergence
    situations during your numerical studies.}\\[0.1cm]
  \textbf{\underline{Response:}} Yes, we have 37 situations where the
  algorithm did not converge in the simulation study (see table
  below). This occurred only for strong overdispersion and small sample
  sizes. We reported these situations in the updated version of the
  paper.
  \begin{table}[H]
    \centering
    \begin{tabular}{lcccc}
      \hline
      & $\phi=-1.6$ & $\phi=-1$ & $\phi=0$ & $\phi=1.8$ \\
      \hline
      $n=50$ &    29 &  0 &   0 &   0 \\
      $n=100$ &   8  &  0 &   0 &   0 \\
      $n=300$ &   0  &  0 &   0 &   0 \\
      $n=1000$ &  0  &  0 &   0 &   0 \\
      \hline
    \end{tabular}
  \end{table}
\item \textit{Numerical derivatives with CMP tend to be very slow and
    prone to errors. It would be good to have a computation times from
    the method for different sample sizes and dispersion
    levels.}\\[0.1cm]
  \textbf{\underline{Response:}} We have computed the times to fit the
  models for the case studies in the paper, they have different
  dispersion levels and sample sizes. However, the Figure below show the
  times to fit in the simulation study. The results show that it took
  more time to fit the models under equidispersion. This result is due
  to choice of starting values and normalizing constant upper bound. We
  set, for all models, the upper bound as 500 and we start the algorithm
  from the true parameters.

  \hspace*{-1cm}
\begin{minipage}{\textwidth}
<<simul-times, fig.width=9, fig.height=3.5>>=
@
\end{minipage}
\vspace{0.1cm}

\item \textit{In Section 5, the simulated counts are ranging from 3 to
    27. This is too limited. Please increase the range.}\\[0.1cm]
  \textbf{\underline{Response}:} The simulation study was performed with
  expectations varying from 3 to 27. Therefore, considering the
  different scenarios ($\phi=-1.6, \phi=-1, \phi=0$ and $\phi=1.8$), we
  cover a wide range of counts. The Figure below shows the variation of
  the counts for $\mu=27$ for the four different scenarios.

@@

\item \textit{From the presentation point of view, focusing more on
    applications would be good as there is not much novelty in the
    methodology.}\\[0.1cm]
  \textbf{\underline{Response:}} In this paper, we propose a
  mean-parametrization of the COM-Poisson model. This parametrization
  introduced $\mu = \lambda^{1/\nu} -(\nu-1)/(2\nu)$, a simple function
  of the original parameters $\lambda$ and $\nu$. We showed that the new
  parameter space has nice properties like i) the orthogonality between
  $\mu$ and $\phi$ (and consequently between $\beta_j$ and $\phi$ in a
  regression setting), leading to better computation and asymptotic
  (normal-based) inference \citep{Ross1970}; and ii) interpretability of
  the $\beta_j$ in terms of the rate ratios since
  $\mu_i\approx \text{E}(Y_i)$. Moreover, we characterized the
  COM-Poisson distribution in terms of the dispersion, zero-inflation
  and heavy-tail indexes that illustrate its applicability to real count
  data. These are the methodological novelties of the paper besides the
  exploration of the results. However, in the updated version of the
  paper, we provided a better discussion of the case studies.
\end{enumerate}

@@

\textbf{\underline{Third paragraph}:} \textit{The authors also hint that
  $var(Y_i)\approx \mu_i \exp(-\phi)$ (see the quadratic error analysis
  leading to Fig. 1). This is curious given that
  $avar(Y_i)=\mu\exp(-\phi) + (\exp(\phi)-1)/(2\exp(2\phi))$ and the
  latter summand converges to $-\infty$ as $\phi\to-\infty$. This latter
  summand is 0 or close to 0 for $\phi=0$ or $\phi>0$. Taken together,
  these last two sentences imply that the $\mu_i\exp(-\phi)$ may be a
  reasonable approximation to the variance when there is equi- or under-
  dispersion, but it may be unreasonable when there is lots of
  overdispersion. It seems that some of the results, simulations, and
  sample analyses reflect this (e.g. Fig 3, Fig 6-8, ``deviance function
  shape under strong overdispersion $\phi=-1.6$ is not as well
  behaved...'', p. 19, etc.). This needs to be addressed.}\\[0.3cm]
\textbf{\underline{Response:}} We thank the referee for exploring this
point. However, we didn't use the approximation of variance
($a\text{Var}(Y_i)$) to reparametrize or fit the COM-Poisson model. To
compute the variances, that are shown in the Figures 3(a) and (b) we
used
$\text{Var}(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y p(y)]^2$.
As highlighted in the Figure 1(b), the approximation is not accurate,
mainly for overdispersion ($\phi<0$). The Figure (a) and (b) below shows
the behavior of $a\text{Var}(Y_i)$ function by fixing $\mu$ and $\phi$,
respectively, illustrating the referees' statement. The $a\text{Var}(Y)$
leads to negative values for variance for small $\mu$ and $\phi<0$ that
shows that the approximation is not reasonable for this region of the
parameter space.

@@

\textbf{\underline{Fifth paragraph}:} \textit{This last point also hints
  that a quasi-Poisson model based on the assumption that
  $var(Y_i)\propto\mu_i$ (see p. 14) may not be reasonable for analyzing
  overdispersed data if the $COMPo$ model truly holds. Is this the
  case?}\\[0.1cm]
\textbf{\underline{Response:}} The quasi-Poisson model is specified by
second-moment assumptions (expectation and variance). As highlighted in
Figure 3(a) (and in referee's previous comment), the mean-variance
relationship for COM-Poisson is linear. Therefore, it can be appropriate
to analyze under-, equi-, and overdispersed data generated according to
the COM-Poisson distribution, using the assumption
$\text{Var}(Y_i)\propto \mu_i$. The advantage of the COM-Poisson
approach is that it corresponds to a fully specified probability model
allowing to compute the likelihood and some useful measures like
deviance, AIC, BIC, LRT, etc.  \vspace{0.4cm}

\textbf{\underline{Sixth paragraph}:} \textit{Consider the example of
  Section 6.1. If I understand matters correctly, it is surprising that
  the $\beta$ estimates under the
  $[COMPo(\lambda_i, \nu), g(m(\lambda_i, \nu)) = \beta^\top x_i]$ model
  are so different from those under the
  $[COMPo(\mu_i, \phi), g(\mu_i) = \beta^\top x_i]$. See the results in
  Table 2, for example. I would think that if the approximation
  $m(\lambda_i, \nu) \approx \mu_i$ is reasonable, which your quadratic
  error results seem to indicate (see Fig 1), then the ML estimates of
  $\beta$ would be very similar. Exactly what models are being fitted
  here? This needs clarification.}\\[0.1cm]
\textbf{\underline{Response:}} Note that the $\lambda_i$ does not
represent the expectation, nor approximately. The errors in Figure 1(a)
is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.  The models
fitted in the case studies are
$[\text{COMPo}(\lambda_i, \nu), \log(\lambda_i) = \beta^\top x_i]$ and
$[\text{COMPo}(\mu_i, \phi), \log(\mu_i) = \beta^\top x_i]$. Therefore,
the $\beta$'s estimates are comparable only if $\phi=0$ (Poisson special
case) otherwise they are on different scales. We make it clear in which
parameter the linear predictor is placed in the current version of the
paper. To understand the relationship of the coefficients, consider
$\log(\lambda_i) = \beta^o$ and $\log(\mu_i) = \beta^*$, so $\beta^*$
expressed in terms of $\beta^o$ is
$\beta^* = \exp(\beta^o-\phi) - (\exp(\phi) - 1)/(2\exp(\phi))$.
\vspace{0.4cm}

@@

\begin{enumerate}
\item \textit{Consider the constraint ``$\mu > 0$'', after equation
    (3.2). What constraints does this impose on the $\lambda$ and
    $\nu$. Is this reasonable?}\\[0.1cm]
  \textbf{\underline{Response:}} The constraint $\mu>0$ implies that
  $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figures below show (a) the
  constraint border in the parameter space; and (b) the expected values
  (obtained by $\sum y p(y)$). The constraint applies to very small
  values of $\lambda$ when $\nu>1$ (underdispersion). This infeasible
  parameter region is related to small expected values (smaller than
  0.1) and underdispersion. Therefore, although this constraint is
  undesirable, it does not prejudice the application of the model. We
  thank the referee for highlighting this point and we have discussed it
  in the updated version.

@@

\item \textit{The Shmueli approximations to the mean and variance hold
    under certain conditions. Are those conditions met in practice? For
    your examples?}\\[0.1cm]
  \textbf{\underline{Response:}} We used only the mean approximation
  given by Shmueli to reparametrize the COM-Poisson model. Moreover, as
  discussed in Figure 1(a), the errors are close to 0 for the parameter
  grid evaluated and present no clear relation with regions gives by
  \cite{Shmueli2005} ($\phi \leq 0$ and
  $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$. The paper proposal is a
  parameterization, so it is always valid (except for the parameter
  constraint imposed by the transformation). Only the interpretation of
  the $\mu$ parameter depends on the approximation accuracy. For
  examples presented in the paper, the parameter estimates (on the
  original parametrization) are summarized below. The approximation is
  good and therefore the $\beta$'s estimates for the Poisson model are
  close to the $\beta$'s estimates for the reparametrized COM-Poisson
  model.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Aug 21 00:57:24 2018
\begin{table}[H]
\centering
\begingroup\small
\begin{tabular}{lrrrrr}
  \toprule
 & & & \multicolumn{3}{c}{$\hat{\lambda}_i$}\\
 \cmidrule(lr){4-6}
Case study & $\hat{\phi}$ & $\hat{\nu}$ & Minimun & Median & Maximum \\
  \midrule
  Cotton experiment   & 1.58 & 4.86 & 3.50 & 8.44 & 9.48 \\
  Soybean experiment  & $-$0.78 & 0.46 &116.87 & 174.98 & 252.29 \\
  Nitrofen experiment & 0.05 & 1.05 & 5.94 & 28.00 & 32.36 \\
   \bottomrule
\end{tabular}
\endgroup
\end{table}
